Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 300\,pc Halo

### 2021-02-12 cluster_paper/ms.tex
update paper

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these halos could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a halo with a length of 300\,pc reported around
  the $\approx$120\,Myr open cluster NGC\,2516.  The classical tidal
  radius of the cluster is $\sim$10\,pc.  Using photometry from Gaia,
  rotation periods from TESS, and lithium measurements from Gaia-ESO
  and GALAH, we find that the halo of NGC\,2516 is real and coeval
  with the core.  Two in three kinematically selected halo members
  25-250\,pc from the cluster center show rotation periods consistent
  with a gyrochronological age of 120\,Myr.  A comparison sample of
  field stars shows no such trend.  The rotation period and lithium
  abundances of the cluster stars are also anti-correlated at fixed
  stellar mass, as is the case with comparable stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.  We highlight implications
  for spectroscopic survey targeting, open cluster dispersal, and
  planet searches around young stars.
\end{abstract}

@@

In the traditional picture of star formation molecular clouds go
gravitationally unstable, and collapse into little knots.  The knots
produce groups of stars close to each other.  The resulting ``open star
clusters'' are, as with any first-order guess, spherical.

How does the cluster evolve?  The answer depends on the total stellar
mass.  The smallest 90\% of star clusters disperse during the ``embedded
phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
by {\it process A} (CITE).  The more massive clusters make it maybe 100
Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
process B}.  Of course, the expected evaporation time should depend on
factors including the mass of the cluster itself, and how many high-mass
(O and B) stars form, since their winds [and maybe supernovae] clear out
most of the cloud, and the initial density of the gas cloud to begin
with.

Assuming that the cluster achieves virial equilibrium, stellar fly-bys
then conspire to segregate the stellar mass distribution within the
cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
assuming equipartition of the specific ``thermal energy'' per star, then
kT = 0.5$mv^2$ at lower masses requires higher velocities).

Identification of the stars that disperse into the galactic field is an
important task for understanding the conditions under which stars and
star clusters form, and for understanding how they subsequently evolve.
For instance, how does the process of radial migration across the
galactic disk affect cluster dispersal?  Did the Sun form in an open
cluster? If so, how massive was its cluster, and is there any hope at
identifying the stars that formed near the Sun?  Qualitatively, the
dispersal of open clusters also provides perhaps the best test case for
the concept of ``chemical tagging'', aka.\ ``galactic archaeology''
(CITE).  If you can't chemically tag stars that are kinematically
associated, why would you be able to do so once they've dispersed
throughout the galactic disk?

Outside of the issue of star formation, identifying the remnant halos of
open clusters is important for a separate project: that of discovering
young transiting planets.  Young transiting planets are hard to find
because young stars are rare (CITE), and often reside in crowded regions
of the sky (CITE).  If the halos of nearby star clusters can be reliably
identified, this could expand the census of nearby sub-Gyr stars by a
factor of 2 or 3 (CITE).  
One path to identify these young stars is to combine photometric
selection criteria with spectroscopic observations (e.g., CITE Zerjal
2019, 2020).
A fortuitous benefit of searching for planets
in cluster environments is that issues with stellar crowding are also
alleviated.

Recent clustering studies using Gaia have begun to report the
identification of structures that could correspond to low-density
halos of stars that may have evaporated from open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  There is evidence that in some
cases, these halos correspond to tidal tails, e.g., in the Hyades
(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
group (CITE Gagne+20).  However, different clustering methods on the
Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
Employing say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space (position or
velocity) unsurprisingly yields open clusters that are roughly
elliptical.  Some unsupervised clustering methods lead to comparable
``roughly elliptical'' results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods such as HDBScan (e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
been found to yield additional structures, particularly in lower
density regions such as the Psc-Eri stream (CITE: Meingast+18,
Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
also more generally around many open clusters
\citep{kounkel_untangling_2019}.  Unsupervised approaches that
incorporate physical constraints like imposing a maximum velocity
dispersion on putative members yield similar results, potentially with
higher purity (Meingast+21).

>>>>

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these halos could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$120\,Myr open cluster NGC\,2516.
  The classical tidal radius of the cluster is $\sim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as is the case with
  comparable stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form in overdensities of hierarchically structured
molecular clouds (CITE).  Over the first $\approx$10\,Myr,
``feedback'' effects including protostellar outflows, photoionization,
radiation pressure, and supernova shocks conspire to disperse the gas
out of the cloud (CITE Krumholz+19).  Since only a small fraction of
the cloud mass is converted into stars ($\approx 1\%$; see Krumholz+19
Figure 10), this gas dispersal drives the majority ($\sim$90\%) of the
stars in the cluster to become gravitationally unbound (e.g., Lada \&
Lada 03).

From 10 to 1000\,Myr, the portion of the cluster that survives gas
dispersal suffers an onslaught that almost always succeeds in
destroying the cluster.  Internal to the cluster, mass loss from
supernovae and AGB winds, and close stellar encounters
(``relaxation'') are important over 10s and 100s of megayears,
respectively.  External to the cluster, collisions with GMCs (Spitzer
1958), and perturbations from the galactic tide in both the radial
(CITE) and vertical (CITE) dimensions further drive stars toward the
``high velocity tails'' of the stellar energy distribution.

% In the traditional picture of star formation molecular clouds go
% gravitationally unstable, and collapse into little knots.  The knots
% produce groups of stars close to each other.  The resulting ``open star
% clusters'' are, as with any first-order guess, spherical.

% How does the cluster evolve?  The answer depends on the total stellar
% mass.  The smallest 90\% of star clusters disperse during the ``embedded
% phase'' (CITE Lada \& Lada 03, Krumholz 19 ARAA).  This is driven mostly
% by {\it process A} (CITE).  The more massive clusters make it maybe 100
% Myr. Maybe a bit longer.  Their evaporation is thought to be driven by
% collisions with molecular clouds (CITE, Spitzer 1958, maybe Sec 7.2 of
% Ryden), and the galactic tide (CITE), and {\it process A}, and {\it
% process B}.  Of course, the expected evaporation time should depend on
% factors including the mass of the cluster itself, and how many high-mass
% (O and B) stars form, since their winds [and maybe supernovae] clear out
% most of the cloud, and the initial density of the gas cloud to begin
% with.
% 
% Assuming that the cluster achieves virial equilibrium, stellar fly-bys
% then conspire to segregate the stellar mass distribution within the
% cluster, evaporating the lowest mass stars soonest (2T+U=0, U=-GM/r,
% assuming equipartition of the specific ``thermal energy'' per star, then
% kT = 0.5$mv^2$ at lower masses requires higher velocities).

Identification of the stars that disperse into the galactic field is an
important aspect of understanding the conditions under which stars and
star clusters form, and for understanding how they subsequently evolve.
For instance, how does the process of radial migration across the
galactic disk affect cluster dispersal?  Did the Sun form in an open
cluster? If so, how massive was its cluster, and is there any hope at
identifying the stars that formed near the Sun?  Qualitatively, the
dispersal of open clusters also provides perhaps the best test case for
the concept of ``chemical tagging'', aka.\ ``galactic archaeology''
(CITE).  If you can't chemically tag stars that are kinematically
associated, why would you be able to do so once they've dispersed
throughout the galactic disk?

Outside of the issue of star formation, identifying the remnant
halos\footnote{ The low-density outer regions of open clusters have
been referred to in general terms as ``halos'' and ``coronae''; we
adopt the former (CITE, CITE). } of open clusters is important for a
separate project: that of discovering young transiting planets.  Young
transiting planets are hard to find because young stars are rare
(e.g., CITE Piskunov 2018), and often reside in crowded regions of the
sky (e.g., CITE ANY OPEN CLUSTER CATALOG).  If the dispersed halos of
nearby star clusters could be reliably identified in a complete
manner, this could expand the census of nearby star clusters by up to
a factor of $10$, based on the expected fraction of stars thought to
be lost during gas dispersal.  Path to identifying these young stars
include photometric criteria (CITE Zari+18), spectroscopic criteria
(CITE Berger+18), and combinations thereof (e.g., CITE Zerjal 2019,
2020, Zhou's work).  A benefit of searching for planets in
dispersed stellar environments is that issues with stellar crowding
are also alleviated.

The identification of low-density (GIVE A NUMBER DENSITY) stellar
associations, already possible before Gaia (e.g.,
Zuckerman+ CITE), has reached a breakthrough pace with the Gaia data.
Some low-density populations (``moving groups'') have no discernable
cores (e.g., USco CITE, Psc Eri CITE, mu Tau CITE).
However other
low-density
halos seem to be centered on previously reported open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  In some
cases, these halos correspond to tidal tails, e.g., in the Hyades
(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
group (CITE Gagne+20). 
However, different clustering methods on the
Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
Employing say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space (position or
velocity) unsurprisingly yields open clusters that are roughly
elliptical.  Some unsupervised clustering methods lead to comparable
``roughly elliptical'' results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods such as HDBScan (e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
been found to yield additional structures, particularly in lower
density regions such as the Psc-Eri stream (CITE: Meingast+18,
Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
also more generally around many open clusters
\citep{kounkel_untangling_2019}.  Unsupervised approaches that
incorporate physical constraints like imposing a maximum velocity
dispersion on putative members yield similar results, potentially with
higher purity (Meingast+21).
